---
title: "PA1_template"
author: "Melanie Fische"
date: "3 Februar 2017"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reproducible Research: Peer Assessment 1

##Loading the file

```{r}
setwd('M:/03_Plan_net_media/13_Insights/02_Intelligence/04_Interne_Projekte/Melanie/Coursera/Kurs 5/Assigment1')
data <- read.csv("activity.csv")
```

### Calculate and report the total number of missing values in the dataset

And filled it with missing data

```{r}
data_na <- na.omit(data)
gesamt_data_na <- aggregate(steps ~ date, data_na, sum)
```

# What is mean total number of steps taken per day?

```{r}
hist(
        gesamt_data_na$steps, 
        main = "total number of steps taken each day", 
        xlab="Steps per day", 
        ylab="Frequency", 
        col = "blue", 
        breaks = 15
)
```

### mean /median

```{r}
meanx <- mean(gesamt_data_na$steps)
print(meanx)

median <- median(gesamt_data_na$steps)
print(median)
```
# What is the average daily activity pattern?

Print a series Plot. At first aggregate the date for series.

```{r}
serien <- aggregate(steps ~ interval, data = data, FUN = function(x) {
        mean(x, na.rm = TRUE)})
library(ggplot2)
ggplot(serien, aes(interval, steps)) + geom_line(colour = "green", lwd = 2) + 
        labs(title = expression("5-minute interval"))
```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
serien[which.max(serien$steps),]
```
# Imputing missing values

1. Calculate the total number.
```{r}
na<-sum(is.na(data))
part<-paste(round(100*(na/nrow(data)), 3), "%")
print(part)
```
2. Devise a strategy for filling in all of the missing values in the dataset...
```{r}
full_data <- data
```
Replace each missing value with the mean value.
```{r}
fuehlenMean <- function(steps, interval) {
        fuehle <- NA
        if (!is.na(steps))
                fuehle <- c(steps)
        else
                fuehle <- (serien[serien$interval==interval, "steps"])
        return(fuehle)
}
full_data$steps <- mapply(fuehlenMean, full_data$steps, full_data$interval)
```
Make a histogram of the total number.
```{r}
total <- tapply(full_data$steps, full_data$date, sum, na.rm=TRUE)
hist(total, main = 'total number of steps taken each day', col = "red")
```
### mean/median of the steps
```{r}
mean1 <- mean(total, na.rm= TRUE)
print(mean1)

median1 <- median(total, na.rm= TRUE)
print(median1)
```
# Are there differences in activity patterns between weekdays and weekends

Create a new weekday/Weekend variable in a df.
```{r}
df_new <- data
df_new$datum <- as.Date(strptime(df_new$date, format="%Y-%m-%d")) 
df_new$tag <-  factor(ifelse(as.POSIXlt(df_new$date)$wday %in% c(0,6), 'weekend', 'weekday'))
```
Aggregate the steps to the different days.
```{r}
week_average_df <- aggregate(full_data$steps, list(interval = data$interval, day = df_new$tag), mean)
names(week_average_df) <- c("interval", "day", "steps")
```
Plot it.
```{r}
ggplot(week_average_df, aes(interval, steps)) + geom_line(color = "purple3", lwd = 1) + 
        xlab("Interval: 5 minute") + ylab("Steps numbers") + facet_grid(day ~ .)
```